Fractional Types

### 2016-12-19 typespp.tex
spell check and clarification about quantifiers

\begin{definition}[Iterated Power of a 1-combinator]
  The $k^{th}$ iterated power of a 1-combinator $p : \tau \iso \tau$, for
  $k \in \Z$ is
  \[
    p^k =
  \begin{cases}
    \idiso & k = 0 \\
    p \odot p^{k - 1} & k > 0 \\
    (!~p) \odot p^{k + 1} & k < 0 \\
  \end{cases}
  \]
\end{definition}

@@

introduces an $\exists$ in the right-to-left direction.

@@

A type $\mathbb{T}_\bullet$ is generally of the form
$\forall\exists{\overline{k}}.[\mathbb{T},v]$ where all the quantifiers are
lifted to the top leaving a non-dependent type and a value in that type. The
combinators in Fig.~\ref{pifracd:comb} also serve as the operational semantics
for the language as they explain how the value currently in focus changes during
evaluation. Given the new combinators, we can modify the the circuit in
Fig.~\ref{fig:zigzag} by adding explicit synchronization primitives. The
evaluation proceeds as follows where we diambiguate the uses of $\forall$ and
$\exists$ for clarity.

@@

\paragraph*{Correspondence with Commutative Semifields.} Computations over
finite types naturally emerge from viewing types as syntax for semiring
elements, semiring identities as type isomorphisms, and justifications for
semiring identities as program transformations and
optimizations~\cite{Carette2016}. This correspondence provides a rich
proof-relevant version of the Curry-Howard correspondence between algebra and
reversible programming languages. The addition of full fractional types to the
mix would enrich the correspondence to commutative semifields, providing a
categorification~\cite{math/9802029} of the non-negative rational numbers in a
computational setting. This correspondence in $\Pi^/$ is still lacking,
however. Given any non-negative rational number, we can form a type whose
cardinality is that number. And yet, our types do not capture the full structure
of the non-negative rational numbers, as these form a commutative
semifield. There are a number of operations and properties which need to be
added to complete this.  Most seem quite straightforward.  We can however single
one out which may not be: what we are missing is a multiplicative inverse for
every type, and not just $\order{p}$. In particular, we would like to form the
type $\iorder{(\iorder{p})}$ and have it be equivalent to $\order{p}$.  As
explained in Sec.~\ref{sec:groupoids}, this would require a futher
generalization of looping and delooping that can be iterated. We leave a
possible extension with a general looping/delooping process to future work.

>>>>

\begin{definition}[Iterated Power of a 1-combinator]
  The $k^{\mbox{th}}$ iterated power of a 1-combinator $p : \tau \iso \tau$, for
  $k \in \Z$ is
  \[
    p^k =
  \begin{cases}
    \idiso & k = 0 \\
    p \odot p^{k - 1} & k > 0 \\
    (!~p) \odot p^{k + 1} & k < 0 \\
  \end{cases}
  \]
\end{definition}

@@

introduces an $\exists$ in the right-to-left direction. A rule
\forall\exists{k}.\mathbb{T_2}_\bullet$ maintains (think of elimination then
introduction) the universal quantifier in the left-to-right direction and
the existential quantifier in the right-to-left direction. We omit the
quantifiers when the sequence of bound variables is empty.

@@

A type $\mathbb{T}_\bullet$ is generally of the form
$\forall\exists{\overline{k}}.[\mathbb{T},v]$ where all the quantifiers are
lifted to the top leaving a non-dependent type and a value in that type. The
combinators in Fig.~\ref{pifracd:comb} also serve as the operational semantics
for the language as they explain how the value currently in focus changes during
evaluation. Given the new combinators, we can modify the the circuit in
Fig.~\ref{fig:zigzag} by adding explicit synchronization primitives. The
evaluation proceeds as follows where we disambiguate the uses of $\forall$ and
$\exists$ for clarity.

@@

\paragraph*{Correspondence with Commutative Semifields.} Computations over
finite types naturally emerge from viewing types as syntax for semiring
elements, semiring identities as type isomorphisms, and justifications for
semiring identities as program transformations and
optimizations~\cite{Carette2016}. This correspondence provides a rich
proof-relevant version of the Curry-Howard correspondence between algebra and
reversible programming languages. The addition of full fractional types to the
mix would enrich the correspondence to commutative semifields, providing a
categorification~\cite{math/9802029} of the non-negative rational numbers in a
computational setting. This correspondence in $\Pi^/$ is still lacking,
however. Given any non-negative rational number, we can form a type whose
cardinality is that number. And yet, our types do not capture the full structure
of the non-negative rational numbers, as these form a commutative
semifield. There are a number of operations and properties which need to be
added to complete this.  Most seem quite straightforward.  We can however single
one out which may not be: what we are missing is a multiplicative inverse for
every type, and not just $\order{p}$. In particular, we would like to form the
type $\iorder{(\iorder{p})}$ and have it be equivalent to $\order{p}$.  As
explained in Sec.~\ref{sec:groupoids}, this would require a further
generalization of looping and delooping that can be iterated. We leave a
possible extension with a general looping/delooping process to future work.
